---
title: "Final Project- Data Demo"
author: "Rohan Dalal-PSTAT 131"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

#DATA MEMO

##Introduction to Data source
I will be choosing COVID data from Covidcast package for my Final Project. 
The covidcast R package, which provides access to the COVIDcast Epidata API published by the Delphi group at Carnegie Mellon University. According to the covidcast R package website,This API provides daily access to a range of COVID-related signals Delphi that builds and maintains, from sources like symptom surveys and medical claims data, and also standard signals that we simply mirror, like confirmed cases and deaths.
([see website here](https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html)) Here is a list of the signals, we can see all the documentation for each one.  This includes information about when the first data points were collected, if the data is available on a daily, or weekly basis, what regions we can call the signal for, and so on.

## Data overview: Fetching/Merging/Prepping data

I plan to choose five signals to predict cases across California counties. 
 Predictor : "visits","admits", "chngVisits" ,"covidChngVisits","gsymptoms" 
and Outcome : "Cases"

- "Cases": Get the number of daily new Covid cases for all the counties in  California,for a given date range (example :from May 2020 to July 2020) by fetching the "US Facts Cases and Deaths" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/usa-facts.html). This will be the Ground Truth(label)

- "visits": Get the daily percentages of doctor visits that are related to Covid in California for a given date range (example :from May 2020 to July 2020)  by fetching the "Doctor Visits" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/doctor-visits.html).

- "admits" :Get the daily hospital admissions for covid diagnosed that are related to Covid in California for a given date range (example :from May 2020 to July 2020)  by fetching the "Doctor Visits" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/hospital-admissions.html).

- "chngVisits": Get the Estimated percentage of outpatient doctor visits primarily about COVID-related symptoms in California for a given date range (example :from May 2020 to July 2020)  by fetching the "Doctor Visits" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/chng.html).

- "covidChngVisits": Get the Estimated percentage of outpatient doctor visits with confirmed COVID-19, based on Change Healthcare claims data in California for a given date range (example :from May 2020 to July 2020)  by fetching the "Doctor Visits" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/chng.html).

- "gsymptoms":Get Sum of Google search volume for anosmia and ageusia related searches in California for a given date range (example :from May 2020 to July 2020)  by fetching the "Doctor Visits" data source (https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/google-symptoms.html).

##Data PreProcessing
- Get the required signals and merge data to create a csv file, clean up and tidy data.
- Observation Count: After collecting needed predictors from the data sources contains about 5428 observations, not all the predictors have missing value but some do.
- Analyze the data fetched for datatype and null/missing values
-- Dealing with missing/NA data :
Method:1-dropping rows with missing values-disadvantage is smaller data set for modeling
Method:2-imputation method-disadvantage it might limit the effectiveness of the model

I am planning to do exploratory analysis to see accuracy and effectivenes compariosn by both methods.


## Motivation/Goal
Goal is to:
-to build predictive models that forecast the future of the pandemic so that we can see one step ahead and prepare accordingly using the past data. 
-to build a predictive model that uses historical COVID cases and related data to forecast the short-term future number of COVID cases in a particular region.

##Project Timeline

-  April 8 -April 14: Load and tidy data 
-  April 14 -April 24: Exploratory analysis and Model selection
-  April 24- May 10: Test and Run models
-  May 10 -May 24 ; work on draft paper
-  May-25 - June 2 : Any edits and finalize paper

##Data collection efforts using covidcast package.

```{r }
#install.packages('covidcast')
library(covidcast)
# Cumulative COVID cases per 100k people on 2020-12-31
df <- covidcast_signal(data_source = "usa-facts", 
                   signal = "confirmed_cumulative_prop",
                   start_day = "2020-12-31", end_day = "2020-12-31")
summary(df)


# This looks at the people who reported COVID-like symptoms from their fb-survey
# from dates 5-1-2020 to 5-7-2020 in all counties 
data <- covidcast_signal("fb-survey", "smoothed_cli", start_day = "2020-05-01",
                         end_day = "2020-05-07")
head(data)


cases <- covidcast_signal(data_source ="usa-facts", "confirmed_incidence_num",
                         start_day = "2020-05-01",
                         end_day = "2020-05-07",
                         geo_type="state", geo_values="ca")
head(cases)
                         



```

```{r }
visits <- covidcast_signal(data_source ="doctor-visits", "smoothed_cli",
                        start_day = "2020-05-01",
                         end_day = "2020-05-07",

                          geo_type="state", geo_values="ca")
head(visits)    

```