---
title: "Homework 2"
author: "PSTAT 131/231"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Linear Regression


### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.
```{r }
abalone <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data", header=FALSE)
colnames(abalone) = c("Sex", "Length", "Diameter", "Height", "Whole", "Shucked","Viscera", "Shell", "Rings") 
head(abalone)

```
```{r }

library(dplyr)
library(rlang)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
tidymodels_prefer()

data<- abalone %>%
  mutate(Age = Rings+1.5)

head(data)

```

```{r }

data %>% 
  ggplot(aes(x = Age)) +
  geom_histogram(bins = 30) +
  theme_bw()

```
Age is normally distributed with slight skewed to the lower end, with a long tail to the right.Most of the abalone is less than 20.

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

```{r }

set.seed(3435)
data_split <- initial_split(data, prop = 0.80,
                                strata = Age)
data_train <- training(data_split)
data_test <- testing(data_split)

head(data_train)

```
*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

```{r }

data_recipe <-  recipe(Age ~ Sex+Length+Diameter+Height+Whole+Shucked+Viscera+Shell, data = data_train) %>% 
  step_dummy(all_nominal_predictors())
data_recipe
summary(data_recipe)

```

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`
    
    
    ```{r }


data_recipe <-  recipe(Age ~ Sex+Length+Diameter+Height+Whole+Shucked+Viscera+Shell, data = data_train) %>% 
  step_dummy(all_nominal_predictors())%>%
  step_interact(terms = ~ Sex:Shucked +Length:Diameter+Shucked:Shell)
data_recipe
summary(data_recipe)

```

3.  center all predictors, and

```{r }
data_recipe <-  recipe(Age ~ Sex+Length+Diameter+Height+Whole+Shucked+Viscera+Shell, data = data_train) %>% 
  step_dummy(all_nominal_predictors())%>%
  step_interact(terms = ~ Sex:Shucked +Length:Diameter+Shucked:Shell)%>%
  step_center(all_predictors())
data_recipe
summary(data_recipe)


```

4.  scale all predictors.
```{r }
data_recipe <-  recipe(Age ~ Sex+Length+Diameter+Height+Whole+Shucked+Viscera+Shell, data = data_train) %>% 
  step_dummy(all_nominal_predictors())%>%
  step_interact(terms = ~ Sex:Shucked +Length:Diameter+Shucked:Shell)%>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
data_recipe
summary(data_recipe)


```

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.


### Question 4

Create and store a linear regression object using the `"lm"` engine.
```{r }
lm_model <- linear_reg() %>% 
  set_engine("lm")%>% 
  set_mode('regression')

```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.
```{r }
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(data_recipe)

```
### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.
```{r }

lm_fit <- fit(lm_wflow, data_train)


data_predict = data.frame("F", 0.50, 0.10, 0.30, 4, 1,2, 1) 
colnames(data_predict) = c("Sex", "Length", "Diameter", "Height", "Whole", "Shucked","Viscera", "Shell")
data_train_res1 <- predict(lm_fit, new_data = data_predict )
data_train_res1
```



### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set includes r square, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the r square value.
```{r }

data_train_res <- predict(lm_fit, new_data = data_train %>% select(-Age))
data_train_res <- bind_cols(data_train_res, data_train%>% select(Age))
data_train_res %>% 
  head()


rmse(data_train_res, truth = Age, estimate = .pred)
data_metrics <- metric_set(rmse, rsq, mae)


data_metrics(data_train_res, truth = Age, 
                estimate = .pred)
```
