---
title: "Homework 3"
author: "PSTAT 131/231"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Classification

For this assignment, we will be working with part of a Kaggle data set that was the subject of a machine learning competition and is often used for practicing ML models. The goal is classification; specifically, to predict which passengers would survive the Titanic shipwreck

Load the data from `data/titanic.csv` into *R* and familiarize yourself with the variables it contains using the codebook.
```{r }
library(dplyr)
titanic <- read.csv("titanic.csv", header=TRUE)

titanic$pclass <- as.factor(titanic$pclass)
titanic$survived <- as.factor(titanic$survived)
titanic$sex <- as.factor(titanic$sex)
titanic <- titanic %>%  mutate(survived=relevel(survived,ref="Yes"))
levels(titanic$survived)

#getwd()
head(titanic)





```

Notice that `survived` and `pclass` should be changed to factors. When changing `survived` to a factor, you may want to reorder the factor so that *"Yes"* is the first level.

Make sure you load the `tidyverse` and `tidymodels`!

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Question 1

Split the data, stratifying on the outcome variable, `survived.`  You should choose the proportions to split the data into. Verify that the training and testing data sets have the appropriate number of observations. Take a look at the training data and note any potential issues, such as missing data.

Why is it a good idea to use stratified sampling for this data?
```{r }
#Analyze the data fetched for datatype and null/missing values
#count of null values in a column
colSums(is.na(titanic))
#position of null values in a column
apply(is.na(titanic), 2, which) 
```

```{r }
set.seed(3435)
library(tidymodels)

titanic_split <- initial_split(titanic, prop = 0.80,
                                strata = survived)
titanic_train <- training(titanic_split)
titanic_test <- testing(titanic_split)

summary(titanic)

str(titanic)
str(titanic_train)
str(titanic_test)
```
Stratifying on the outcome variable, `survived.` provides better coverage of the population isrepresented in the sampling.

### Question 2

Using the **training** data set, explore/describe the distribution of the outcome variable `survived`.

```{r }

summary(titanic_train)

```
Total observations in the Train data set = 712 , where 273 survived and 439 did not. We do have null/missing values in some predictors.not all. 
below is the visualization of training data for actual outcome.

```{r }
titanic_train %>% 
  ggplot(aes(x =survived)) +
  geom_bar()


```



### Question 3

Using the **training** data set, create a correlation matrix of all continuous variables. Create a visualization of the matrix, and describe any patterns you see. Are any predictors correlated with each other? Which ones, and in which direction?


```{r }

library(discrim)
library(poissonreg)
library(corrr)
library(klaR) # for naive bayes
tidymodels_prefer()

data<-titanic_train
data <- data %>%
        select(where(is.numeric))
cor_data <- data %>%
  correlate()
rplot(cor_data)

```
```{r }
cor_data %>%
  stretch() %>%
  ggplot(aes(x, y, fill = r)) +
  geom_tile() +
  geom_text(aes(label = as.character(fashion(r))))
```
Notice that the upper and lower triangles of the matrix are identical; that's a common feature of correlation matrices. The grey squares represent the variances of the variables. Again, we see that only sib_sp(# of siblings / spouses aboard the Titanic) and parch(# of parents / children aboard the Titanic) have much of any correlation with each other, and it's only about 0.45

### Question 4


Using the **training** data, create a recipe predicting the outcome variable `survived`. Include the following predictors: ticket class, sex, age, number of siblings or spouses aboard, number of parents or children aboard, and passenger fare.

Recall that there were missing values for `age`. To deal with this, add an imputation step using `step_impute_linear()`. Next, use `step_dummy()` to **dummy** encode categorical predictors. Finally, include interactions between:

-   Sex and passenger fare, and
-   Age and passenger fare.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r }

data_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, data = titanic_train)%>%
  step_impute_linear(age)%>%
  step_dummy(all_nominal_predictors())%>%
  step_interact(terms = ~ Sex:fare + age:fare)
data_recipe
summary(data_recipe)

```

### Question 5

Specify a **logistic regression** model for classification using the `"glm"` engine. Then create a workflow. Add your model and the appropriate recipe. Finally, use `fit()` to apply your workflow to the **training** data.

***Hint: Make sure to store the results of `fit()`. You'll need them later on.***

```{r }

log_reg <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

log_wkflow <- workflow() %>% 
  add_model(log_reg) %>% 
  add_recipe(data_recipe)

log_fit <- fit(log_wkflow, titanic_train)

log_fit %>% 
  tidy()


```

### Question 6

**Repeat Question 5**, but this time specify a linear discriminant analysis model for classification using the `"MASS"` engine.

```{r }


lda_mod <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")

lda_wkflow <- workflow() %>% 
  add_model(lda_mod) %>% 
  add_recipe(data_recipe)

lda_fit <- fit(lda_wkflow, titanic_train)

lda_fit 

```

### Question 7

**Repeat Question 5**, but this time specify a quadratic discriminant analysis model for classification using the `"MASS"` engine.

```{r }

qda_mod <- discrim_quad() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")

qda_wkflow <- workflow() %>% 
  add_model(qda_mod) %>% 
  add_recipe(data_recipe)

qda_fit <- fit(qda_wkflow, titanic_train)

```

### Question 8

**Repeat Question 5**, but this time specify a naive Bayes model for classification using the `"klaR"` engine. Set the `usekernel` argument to `FALSE`.
```{r }

nb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

nb_wkflow <- workflow() %>% 
  add_model(nb_mod) %>% 
  add_recipe(data_recipe)

nb_fit <- fit(nb_wkflow, titanic_train)

```

### Question 9

Now you've fit four different models to your training data.

Use `predict()` and `bind_cols()` to generate predictions using each of these 4 models and your **training** data. Then use the *accuracy* metric to assess the performance of each of the four models.

Which model achieved the highest accuracy on the training data?

```{r }
#Method1
log_res<-predict(log_fit, new_data = titanic_train, type = "prob")
log_res <- bind_cols(log_res, titanic_train%>% select(survived))

augment(log_fit, new_data = titanic_train) %>%
  conf_mat(truth = survived, estimate = .pred_class)
log_reg_acc <- augment(log_fit, new_data = titanic_train) %>%
  accuracy(truth = survived, estimate = .pred_class)
log_reg_acc

#Method2
lda_res<-predict(lda_fit, new_data = titanic_train, type = "prob")
lda_res <- bind_cols(lda_res, titanic_train%>% select(survived))
augment(lda_fit, new_data = titanic_train) %>%
  conf_mat(truth = survived, estimate = .pred_class) 
lda_acc <- augment(lda_fit, new_data = titanic_train) %>%
  accuracy(truth = survived, estimate = .pred_class)
lda_acc

#Method3
qda_res<-predict(qda_fit, new_data = titanic_train, type = "prob")
qda_res <- bind_cols(qda_res, titanic_train%>% select(survived))
augment(qda_fit, new_data = titanic_train) %>%
  conf_mat(truth = survived, estimate = .pred_class) 
qda_acc <- augment(qda_fit, new_data = titanic_train) %>%
  accuracy(truth = survived, estimate = .pred_class)
qda_acc

#Method4
nb_res<-predict(nb_fit, new_data = titanic_train, type = "prob")
nb_res <- bind_cols(nb_res, titanic_train%>% select(survived))
augment(nb_fit, new_data = titanic_train) %>%
  conf_mat(truth = survived, estimate = .pred_class) 
nb_acc <- augment(nb_fit, new_data = titanic_train) %>%
  accuracy(truth = survived, estimate = .pred_class)
nb_acc

accuracies <- c(log_reg_acc$.estimate, lda_acc$.estimate, 
                nb_acc$.estimate, qda_acc$.estimate)
models <- c("Logistic Regression", "LDA", "Naive Bayes", "QDA")
results <- tibble(accuracies = accuracies, models = models)
results %>% 
  arrange(-accuracies)

```

### Question 10

Fit the model with the highest training accuracy to the **testing** data. Report the accuracy of the model on the **testing** data.

Again using the **testing** data, create a confusion matrix and visualize it. Plot an ROC curve and calculate the area under it (AUC).

How did the model perform? Compare its training and testing accuracies. If the values differ, why do you think this is so?
```{r }
predict(qda_fit, new_data = titanic_test, type = "prob")
augment(qda_fit, new_data = titanic_test) %>%
  conf_mat(truth = survived, estimate = .pred_class) 

multi_metric <- metric_set(accuracy, sensitivity, specificity)

augment(qda_fit, new_data = titanic_test) %>%
  multi_metric(truth = survived, estimate = .pred_class)


```
```{r }

augment(qda_fit, new_data = titanic_test) %>%
  conf_mat(truth = survived, estimate = .pred_class) %>%
  autoplot(type = "heatmap")


```
```{r }

augment(qda_fit, new_data = titanic_test) %>%
  roc_curve(survived, .pred_No) %>%
  autoplot()


```

```{r }
qda_train_acc <- augment(qda_fit, new_data = titanic_train) %>%
  accuracy(truth = survived, estimate = .pred_class)

qda_test_acc <- augment(qda_fit, new_data = titanic_test) %>%
  accuracy(truth = survived, estimate = .pred_class)

accuracies <- c(qda_train_acc$.estimate, qda_test_acc$.estimate )
models <- c("Train Data Accuracy", "Test Data Accuracy")
results <- tibble(accuracies = accuracies, models = models)
results %>% 
  arrange(-accuracies)



```

QDA model perfrom really good. Test data accuracy(79%) is very close to the training data accuracy(81%).